---
title: "STA302"
output: html_document
---

The [R plugin](https://www.jetbrains.com/help/pycharm/r-plugin-support.html) for IntelliJ-based IDEs provides
handy capabilities to work with the [R Markdown](https://www.jetbrains.com/help/pycharm/r-markdown.html) files.
To [add](https://www.jetbrains.com/help/pycharm/r-markdown.html#add-code-chunk) a new R chunk,
```{r}
# install.packages("GGally")
# install.packages("car")
library(tidyverse)
library(car)
library(GGally)
library(caret)
# contains:
# ggplot2: For data visualization.
# dplyr: For data manipulation.
# tidyr: For data tidying.
# readr: For data import.
# purrr: For functional programming.
# tibble: For tibbles, a modern re-imagining of data frames.
# stringr: For string manipulation.
# forcats: For working with factors.
#
```
```{r}
data <- read_csv("savant_data.csv",
                 col_select = c(
                         "pitch_type",
                         "release_speed",
                         # "release_pos_x", "release_pos_z",
                         # "release_extension",
                         "release_spin_rate", "pfx_x", "pfx_z", "p_throws",
                         # "spin_axis",
                         "delta_run_exp"
                    )
                 )
names(data)[1] <- "pitch_type"

```
```{r}
# Create dummy variables for p_throws and pitch_type if not already present
data <- data %>%
    mutate(pitch_type = case_when(
           pitch_type %in% c("FF", "SI", "FC") ~ "fastball",
           pitch_type %in% c("SL", "ST", "SV", "CU", "KC", "CS", "KN") ~ "breaking",
           pitch_type %in% c("CH", "FS", "FO", "SC") ~ "offspeed",
           TRUE ~ "other"
          )) %>%
    filter(pitch_type != "other")

data <- na.omit(data)
```
```{r}
dummies <- dummyVars(~ p_throws + pitch_type, data = data)
one_hot_encoded_data <- predict(dummies, newdata = data)

one_hot_encoded_data <- as.data.frame(one_hot_encoded_data)
X <- data[, !names(data) %in% c("p_throws", "pitch_type", "delta_run_exp")]
X <- cbind(X, one_hot_encoded_data)
X <- X[, !names(X) %in% c("p_throwsR", "pitch_typefastball")]

y <- data$delta_run_exp
```
position the caret at any line or the code chunk, then click "+".

The code chunk appears:
```{r}
# ggpairs(data)
```

```{r}
pitch_proportions <- data %>%
  group_by(pitch_type) %>%
  summarise(count = n()) %>%
  mutate(proportion = count / sum(count))

# Plot the proportions
ggplot(pitch_proportions, aes(x = pitch_type, y = proportion, fill = pitch_type)) +
  geom_bar(stat = "identity") +
  labs(title = "Proportions of Each Pitch Type", x = "Pitch Type", y = "Proportion") +
  theme_minimal()
```
```{r}

```
```{r}

```
```{r}

```
```{r}

```
```{r}
model <- lm(y ~ ., data = X)
summary(model)
```
```{r}
fitted <- fitted(model)
resid <- resid(model)

plot(resid ~ fitted, xlab = "Fitted values", ylab = "Residuals", main = "Residuals vs Fitted")
```
```{r}
qqnorm(resid)
qqline(resid)
```
```{r}
plot(resid ~ X$release_speed, xlab = "Release Speed", ylab = "Residuals", main = "Residuals vs Release Speed")
plot(resid ~ X$pfx_x, xlab = "PFX X", ylab = "Residuals", main = "Residuals vs PFX X")
plot(resid ~ X$pfx_z, xlab = "PFX Z", ylab = "Residuals", main = "Residuals vs PFX Z")
plot(resid ~ X$release_spin_rate, xlab = "Release Spin Rate", ylab = "Residuals", main = "Residuals vs Release Spin Rate")

boxplot(resid ~ X$p_throwsL, xlab = "p_throwsL", ylab = "Residuals", main = "Residuals vs p_throwsL")
boxplot(resid ~ X$pitch_typebreaking, xlab = "pitch_typebreaking", ylab = "Residuals", main = "Residuals vs pitch_typebreaking")
boxplot(resid ~ X$pitch_typeoffspeed, xlab = "pitch_typeoffspeed", ylab = "Residuals", main = "Residuals vs pitch_typeoffspeed")
```

Now, click the **Run** button on the chunk toolbar to [execute](https://www.jetbrains.com/help/pycharm/r-markdown.html#run-r-code) the chunk code. The result should be placed under the chunk.
```{r}
plot(fitted ~ y, xlab = "Fitted values", ylab = "Actual values", main = "Fitted vs Actual")
```
Click the **Knit and Open Document** to build and preview an output.kkjk