---
title: "STA302 Assignment"
author: "Louis Tan, Maverick Luke, Tom Kim"
date: "2024"
geometry: margin=0.5in
output: pdf_document
---

## Outline

-   **Task 0:** Setup
-   **Task 1:** Data Cleaning
-   **Task 2:** Initial Linear Model
-   **Task 3:** Initial Model Diagnostics and Analysis
-   **Task 4:** Update Linear Model
-   **Task 5:** Predictions
-   **Task 6:** Final Diagnostics and Analysis

## Task 0: Setup

```{r}
# Load appropriate libraries
library(ggplot2)
library(dplyr)
library(mgcv)
library(car)

# Load the dataset 
# (Uncomment, load, comment)
# source("HowToSavantData.R", echo = TRUE, print.eval = TRUE)
```

## Task 1: Data Cleaning

-   Filter data
-   Remove incomplete data points

```{r}
# Filter data for right-handed pitchers (RHP) and four-seam fastballs (FF)
# Filter for all swings and calculate whiff rates
RHP_FB <- savant_data |>
  filter(p_throws == "R", pitch_type == "FF",
         description %in% c("hit_into_play", "swinging_strike", "foul", "foul_tip")) |>
  mutate(whiff = if_else(description == "swinging_strike", 1, 0))
```

We isolate the data to right-handed pitchers, as left-handed pitchers fundamentally perform differently (nt. *Platoon effect*). We also isolate to four-seam fastballs, as each pitch has its own distinct performance characteristics.

*Whiff rate* is the percentage of swing strikes per swing. In baseball, it is the most important metric for measuring how *powerful* (i.e. effective) the pitch is in inducing strikeouts (i.e. preventing runs).

```{r}
# Remove incomplete data points (NA's)

RHP_FB0 <- RHP_FB |>
  filter(
    !is.na(stand), 
    !is.na(release_speed),
    !is.na(release_spin_rate), 
    !is.na(release_extension),
    !is.na(release_pos_z), 
    !is.na(release_pos_x),
    !is.na(plate_z), 
    !is.na(plate_x),
    !is.na(ax), 
    !is.na(az),
    !is.na(pfx_x), 
    !is.na(pfx_z)
  ) |>
  group_by(player_name, game_year, stand) |>
  summarise(release_speed = mean(release_speed),
            release_spin_rate = mean(release_spin_rate),
            release_extension = mean(release_extension),
            release_pos_x = mean(release_pos_x),
            release_pos_z = mean(release_pos_z),
            plate_z = mean(plate_z),
            plate_x = mean(plate_x),
            ax = mean(ax),
            az = mean(az),
            pfx_x = mean(pfx_x),
            pfx_z = mean(pfx_z),
            whiff = mean(whiff),
            count = n(),
            .groups="drop") |>
     filter(162 < count & release_pos_z > 5)
```



We group by *player_name* and *game_year* as this classification leads to the most self-consistency within samples. *stand* is also included as it is a categorical variable (see below).

## Task 2: Initial Linear Model

-   Response Variable:
    -   *whiff*: percentage of swinging_strike of all swings
-   Predictor Variable(s):
    -   *stand*: if batter is left/right handed (categorical)
    -   *release_speed*: how fast the pitch is
    -   *release_spin_rate*: how fast the ball spins (in RPM)
    -   *release_extension*: how far the release point is from the rubber
    -   *release_pos_z*: how high the release point is
    -   *release_pos_x*: how far horizontally the release point is from the center
    -   *plate_z*: vertical location of the pitch (where it ends)
    -   *plate_x*: horizontal location of the pitch (where it ends)
    -   *az*: vertical acceleration of the pitch
    -   *ax*: horizontal acceleration of the pitch
    -   *pfx_x*: horizontal movement of the pitch
    -   *pfx_z*: vertical movement of the pitch
-   Interaction Term(s):
    -   *stand : release_pos_x*: being on left/right side of the plate affects the horizontal angle
    -   *stand : plate_x*: to a left-handed batter, higher plate_x means the pitch location is more relatively inside (opposite for right-handers)
    -   *stand : ax*: similar to above, horizontal acceleration is relatively opposite to left-handers

We decide to include *stand* as both a predictor variable and as part of interaction, as left-handed hitters tend to swing differently than right-handed hitters.

For the linear model, we weigh each data point by the number of throws; player_name—year combination with count=100 is considered to be worth 100 data points.

```{r}
# Fit the initial linear model
fb_model <- lm(whiff ~ 
                 # Predictors
                 stand + release_speed + release_spin_rate +
                 release_extension + release_pos_z + 
                 release_pos_x + plate_x + plate_z + 
                 ax + az + pfx_x + pfx_z + 
                 # Interaction terms
                 stand:release_pos_x + stand:ax + stand:plate_x,
               data = RHP_FB0, weights = count)
```

## Data Visualization

```{r}
# Response variable distribution
p_response <- ggplot(data = data.frame(responses = responses), aes(x = responses)) +
  geom_histogram(bins = 30, fill = "#3498db", color = "white", alpha = 0.7) +
  labs(title = "Distribution of Response Variable",
       x = "Response",
       y = "Count") +
  theme
# We can either show the plot immediately using print or save it as an image
print(p_response)
# ggsave("response_distribution.jpg", 
#        p_response, 
#        width = 8, 
#        height = 6, 
#        dpi = 300)
```

```{r}
e_hat <- resid(fb_model)
y_hat <- fitted(fb_model)
```

```{r}
# Primary theme for our plots
theme <- theme_minimal() +
  theme(
    plot.title = element_text(size = 15, face = "bold", margin = margin(b = 10)),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 11),
    plot.margin = margin(10, 20, 10, 10),
    panel.grid.major = element_line(color = "gray90", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white", color = NA)
  )

# Function to create a standardized residual plot
create_residual_plot <- function(data, x_var, x_lab, title) {
  ggplot(data, aes(x = {{x_var}}, y = e_hat)) +
    geom_point(alpha = 0.6, color = "#2c3e50", size = 1.5) +  
    geom_smooth(method = "loess", se = TRUE, color = "#e74c3c", fill = "#ffcdd2", alpha = 0.2) +  
    labs(title = title, 
         x = x_lab, 
         y = "Residuals") +
    theme +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.3)
}

# Residuals vs Predictors
p_resid_rs <- create_residual_plot(RHP_FB0, release_speed, "Release Speed", "Residuals vs Release Speed")
p_resid_rsr <- create_residual_plot(RHP_FB0, release_spin_rate, "Release Spin Rate", "Residuals vs Release Spin Rate")
p_resid_re <- create_residual_plot(RHP_FB0, release_extension, "Release Extension", "Residuals vs Release Extension")
p_resid_rpz <- create_residual_plot(RHP_FB0, release_pos_z, "Release Pos Z", "Residuals vs Release Pos Z")
p_resid_rpx <- create_residual_plot(RHP_FB0, release_pos_x, "Release Pos X", "Residuals vs Release Pos X")
p_resid_pz <- create_residual_plot(RHP_FB0, plate_z, "Plate Z", "Residuals vs Plate Z")
p_resid_px <- create_residual_plot(RHP_FB0, plate_x, "Plate X", "Residuals vs Plate X")
p_resid_az <- create_residual_plot(RHP_FB0, az, "Az", "Residuals vs Az")
p_resid_ax <- create_residual_plot(RHP_FB0, ax, "Ax", "Residuals vs Ax")
p_resid_pfx_z <- create_residual_plot(RHP_FB0, pfx_z, "Pfx Z", "Residuals vs Pfx Z")
p_resid_pfx_x <- create_residual_plot(RHP_FB0, pfx_x, "Pfx X", "Residuals vs Pfx X")

# Special case for boxplot
p_resid_s <- ggplot(data = RHP_FB0, aes(x = stand, y = e_hat)) +
  geom_boxplot(fill = "#3498db", alpha = 0.6, outlier.color = "#e74c3c", outlier.alpha = 0.6) +
  labs(title = "Residuals vs Stand", 
       x = "Stand", 
       y = "Residuals") +
  theme +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.3)

combined_plot <- arrangeGrob(
  p_resid_rs, p_resid_rsr, p_resid_re, p_resid_rpz, 
  p_resid_rpx, p_resid_pz, p_resid_px, p_resid_az, 
  p_resid_ax, p_resid_pfx_z, p_resid_pfx_x, p_resid_s, 
  ncol = 4,
  top = textGrob("Residual Analysis Plots", 
                 gp = gpar(fontsize = 20, fontface = "bold"))
)

# Saved as residual_plots.jpg in the working directory
ggsave("residual_plots.jpg", 
      combined_plot, 
      width = 15, 
      height = 12, 
      dpi = 300)

```

```{r}
# Residuals vs fitted values
p_resid_fitted <- ggplot(data = data.frame(fitted = y_hat, residuals = e_hat), 
                        aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.6, color = "#2c3e50", size = 1.5) +
  geom_smooth(method = "loess", se = TRUE, color = "#e74c3c", fill = "#ffcdd2", alpha = 0.2) +
  labs(title = "Residuals vs Fitted Values",
       x = "Fitted Values",
       y = "Residuals") +
  theme +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.3)

print(p_resid_fitted)
# ggsave("residuals_vs_fitted.jpg", 
#        p_resid_fitted, 
#        width = 8, 
#        height = 6, 
#        dpi = 300)

# Normal Q-Q plot
p_qq <- ggplot(data.frame(residuals = e_hat), aes(sample = residuals)) +
  stat_qq(alpha = 0.6, color = "#2c3e50", size = 1.5) +
  stat_qq_line(color = "#e74c3c") +
  labs(title = "Normal Q-Q Plot",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles") +
  theme

print(p_qq)

# gsave("qq_plot.jpg", 
#        p_qq, 
#        width = 8, 
#        height = 6, 
#        dpi = 300)

predictors_df <- RHP_FB0[, c("release_speed", "release_spin_rate", "release_extension", 
                            "release_pos_z", "release_pos_x", "plate_z", "plate_x",
                            "az", "ax", "pfx_z", "pfx_x")]

p_pairs <- ggpairs(predictors_df,
                   lower = list(continuous = wrap("points", alpha = 0.3, size = 0.5, color = "#2c3e50")),
                   diag = list(continuous = wrap("densityDiag", fill = "#3498db", alpha = 0.5)),
                   upper = list(continuous = wrap("cor", size = 3))) +
  theme

print(p_pairs)
# ggsave("pairs_plot.jpg", 
#        p_pairs, 
#        width = 20, 
#        height = 20, 
#        dpi = 300)


```

## Variable transformation
```{r}
summary(powerTransform(cbind(RHP_FB0[,4:15])))
```

```{r}
# Transform our response variable and do another filter for release pos z (due to special case of a person's pitch specialty of throwing ball as low as possible)
RHP_FB0 <- RHP_FB0 |> 
        mutate(whiff = whiff^(1/3)) |> 
        filter(release_pos_z > 5)

# Rebuild our model with transformed data
fb_model <- lm(whiff ~ 
                 # Predictors
                 stand + release_speed + release_spin_rate +
                 release_extension + release_pos_z + 
                 release_pos_x + plate_x + plate_z + 
                 ax + az + pfx_x + pfx_z + 
                 # Interaction terms
                 stand:release_pos_x + stand:ax + stand:plate_x,
               data = RHP_FB0, weights = count)

e_hat <- resid(fb_model)
y_hat <- fitted(fb_model)
```

```{r}
# QQ-plot for model with transformed data
p_qq <- ggplot(data.frame(residuals = e_hat), aes(sample = residuals)) +
  stat_qq(alpha = 0.6, color = "#2c3e50", size = 1.5) +
  stat_qq_line(color = "#e74c3c") +
  labs(title = "Normal Q-Q Plot",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles") +
  theme

print(p_qq)

# gsave("qq_plot.jpg", 
#        p_qq, 
#        width = 8, 
#        height = 6, 
#        dpi = 300)

```

## Task 3: Initial Model Diagnostics and Analysis

-   Analyse coefficients, correlation, and statistical significance
-   Create condition 2 plot
-   Create Residual plots
-   Create Box Plots
-   Perform Partial F tests
-   Perform VIF test

```{r}
# Analyse coefficients, correlation, and statistical significance
summary(fb_model)
```

We immediately see that the *Adjusted R-squared* is 0.2233, and the *p-value* is \<2.2e-16, which is promising.

We see that some predictors had high p-values—we delve into this later.

```{r}
# Create condition 2 plot
pairs(RHP_FB0[,4:15])
```

Description

Description

Description

Description

Description
```{r}
```
```{r}
di <- cooks.distance(fb_model)
cutoff_di <- qf(0.5, 16, 1543)
print(which(di > cutoff_di))

ri <- rstandard(fb_model)
print(which(ri > 4 | ri < -4))

hii <- hatvalues(fb_model)
cutoff_hii <- 2 * (16) / 1559
which(hii > cutoff_hii)

dffits <- dffits(fb_model)
cutoff_dffits <- 2 * sqrt(16 / 1559)
which(abs(dffits) > cutoff_dffits)

dfbetas <- dfbetas(fb_model)
cutoff_dfbetas <- 7 / sqrt(1559)
which(abs(dfbetas[,5]) > cutoff_dfbetas)
print(RHP_FB0[RHP_FB0$release_pos_z < 6,])
```


```{r}
```


```{r}
hist(RHP_FB0$release_speed)
hist(RHP_FB0$release_spin_rate)
hist(RHP_FB0$release_extension)
hist(RHP_FB0$release_pos_x)
hist(RHP_FB0$release_pos_z)
hist(RHP_FB0$ax)
hist(RHP_FB0$az)
hist(RHP_FB0$whiff)
```

```{r}
# Create Residual plots

# Residual vs Fitted plot
y_hat <- fitted(fb_model)
e_hat <- resid(fb_model)
plot(x = y_hat, y = e_hat, main = "Residual vs Fitted", xlab = "Fitted", ylab = "Residuals")

```

We immediately notice the 3 distinct lines in the residual plot.

To this, we hypothesize the following:

-   The whiff variable is derived as the proportion of swinging strikes (swinging_strike) within grouped data (group_by). Since whiff is calculated as the mean of a binary variable, it can only take on discrete values when grouped over small count sizes.
-   For instance, if there are only 3 pitches in a group, whiff can only take values of 0, 1/3, 2/3, 1. This creates "discrete jumps" in the residuals and leads to the horizontal bands in the plot.

```{r}
# Temporarily filter data points
par(mfrow = c(1,3))

# Count >= 4
RHP_FB0_temp <- RHP_FB0 |>
  filter(count > 3)
fb_model_temp <- lm(whiff ~ 
                 stand + release_speed + release_spin_rate +
                 release_extension + release_pos_z + 
                 release_pos_x + plate_x + plate_z + 
                 ax + az + pfx_x + pfx_z + 
                 stand:release_pos_x + stand:ax + stand:plate_x,
               data = RHP_FB0_temp, weights = count)
y_hat_temp <- fitted(fb_model_temp)
e_hat_temp <- resid(fb_model_temp)
plot(x = y_hat_temp, y = e_hat_temp, main = "Residual vs Fitted", xlab = "Fitted")

# Count >= 16
RHP_FB0_temp <- RHP_FB0 |>
  filter(count >= 16)
fb_model_temp <- lm(whiff ~ 
                 stand + release_speed + release_spin_rate +
                 release_extension + release_pos_z + 
                 release_pos_x + plate_x + plate_z + 
                 ax + az + pfx_x + pfx_z + 
                 stand:release_pos_x + stand:ax + stand:plate_x,
               data = RHP_FB0_temp, weights = count)
y_hat_temp <- fitted(fb_model_temp)
e_hat_temp <- resid(fb_model_temp)
plot(x = y_hat_temp, y = e_hat_temp, main = "Residual vs Fitted", xlab = "Fitted")

# Count >= 64
RHP_FB0_temp <- RHP_FB0 |>
  filter(count >= 64)
fb_model_temp <- lm(whiff ~ 
                 stand + release_speed + release_spin_rate +
                 release_extension + release_pos_z + 
                 release_pos_x + plate_x + plate_z + 
                 ax + az + pfx_x + pfx_z + 
                 stand:release_pos_x + stand:ax + stand:plate_x,
               data = RHP_FB0_temp, weights = count)
y_hat_temp <- fitted(fb_model_temp)
e_hat_temp <- resid(fb_model_temp)
plot(x = y_hat_temp, y = e_hat_temp, main = "Residual vs Fitted", xlab = "Fitted")


# Cleanup
rm(fb_model_temp)
rm(RHP_FB0_temp)

```

After filtering small count sizes, we observe no lines.

We also observe no bias in the residuals, and no distinct skew/patterns.

```{r}
# Residual vs Predictor plots for each predictor variable
plot(RHP_FB0$release_speed, e_hat, main="Residuals vs Release Speed", xlab="Release Speed", ylab="Residuals")
plot(RHP_FB0$release_spin_rate, e_hat, main="Residuals vs Release Spin Rate", xlab="Release Spin Rate", ylab="Residuals")
plot(RHP_FB0$release_extension, e_hat, main="Residuals vs Release Extension", xlab="Release Extension", ylab="Residuals")
plot(RHP_FB0$release_pos_z, e_hat, main="Residuals vs Release Pos Z", xlab="Release Pos Z", ylab="Residuals")
plot(RHP_FB0$release_pos_x, e_hat, main="Residuals vs Release Pos X", xlab="Release Pos X", ylab="Residuals")
plot(RHP_FB0$plate_z, e_hat, main="Residuals vs Plate Z", xlab="Plate Z", ylab="Residuals")
plot(RHP_FB0$plate_x, e_hat, main="Residuals vs Plate X", xlab="Plate X", ylab="Residuals")
plot(RHP_FB0$ax, e_hat, main="Residuals vs Ax", xlab="Ax", ylab="Residuals")
plot(RHP_FB0$az, e_hat, main="Residuals vs Az", xlab="Az", ylab="Residuals")
plot(RHP_FB0$pfx_x, e_hat, main="Residuals vs Pfx X", xlab="Pfx X", ylab="Residuals")
plot(RHP_FB0$pfx_z, e_hat, main="Residuals vs Pfx Z", xlab="Pfx Z", ylab="Residuals")
```

Description

Description

Description

Description

Description
```{r}
```


```{r}
# Create Box Plots
# boxplot(e_hat ~ RHP_FB0$release_speed, main="Residuals vs Release Speed", xlab="Release Speed", ylab="Residuals")
# boxplot(e_hat ~ RHP_FB0$release_spin_rate, main="Residuals vs Release Spin Rate", xlab="Release Spin Rate", ylab="Residuals")
# boxplot(e_hat ~ RHP_FB0$release_extension, main="Residuals vs Release Extension", xlab="Release Extension", ylab="Residuals")
# boxplot(e_hat ~ RHP_FB0$release_pos_z, main="Residuals vs Release Pos Z", xlab="Release Pos Z", ylab="Residuals")
# boxplot(e_hat ~ RHP_FB0$release_pos_x, main="Residuals vs Release Pos X", xlab="Release Pos X", ylab="Residuals")
# boxplot(e_hat ~ RHP_FB0$plate_z, main="Residuals vs Plate Z", xlab="Plate Z", ylab="Residuals")
# boxplot(e_hat ~ RHP_FB0$plate_x, main="Residuals vs Plate X", xlab="Plate X", ylab="Residuals")
# boxplot(e_hat ~ RHP_FB0$ax, main="Residuals vs Ax", xlab="Ax", ylab="Residuals")
# boxplot(e_hat ~ RHP_FB0$az, main="Residuals vs Az", xlab="Az", ylab="Residuals")
# boxplot(e_hat ~ RHP_FB0$pfx_x, main="Residuals vs Pfx X", xlab="Pfx X", ylab="Residuals")
# boxplot(e_hat ~ RHP_FB0$pfx_z, main="Residuals vs Pfx Z", xlab="Pfx Z", ylab="Residuals")
```

Description

Description

Description

Description

Description

```{r}
# Create QQ plot
qqnorm(e_hat)
qqline(e_hat)
```

Description

Description

Description

Description

Description

Earlier, we noticed that some predictors appeared to be statistically insignificant—*release_pos_x* and *ax*.

Before outright removing these predictors, we suspect that this may be due to multicollinearity. To demonstrate this, we perform the VIF test.

```{r}
# Perform VIF test
vif(fb_model)
```

Here, we see a high (\>10) inflation factor(s) for *stand*, *release_speed*, *az*, *ax*, *pfx_x*, *pfx_z*, *stand:release_pos_x*, *plate_x*. This hints that these variables may be collinear with *release_pos_x* and *ax*. However, *release_pos_x* did not have a high inflation factor. Also, we have observed an exceptionally high p-value of *release_pos_x*—0.881024—earlier. *release_pos_x* also had a very small coefficient estimate. These may be an implication that it is an insignificant variable. We perform an ANOVA test to verify this.

```{r}
# Fit temporary model without release_pos_x
fb_model_temp <- lm(whiff ~ 
                 stand + release_speed + release_spin_rate +
                 release_extension + release_pos_z + 
                 plate_x + plate_z + 
                 ax + az + pfx_x + pfx_z + 
                 stand:ax + stand:plate_x,
               data = RHP_FB0, weights = count)

# Perform ANOVA test
anova(fb_model_temp, fb_model)
```

Removing *release_pos_x* results in a statistically significant loss in explanatory capacity. Thus, we hypothesize the following: although *release_pos_x* is a significant predictor variable, the interaction term *stand:release_pos_x* overwhelms it, and is collinear to *stand* (as demonstrated by the VIF test). The reasoning is that pitchers vary their release point depending on the handedness of the hitter to maximize the horizontal effects against the hitters.

Hence, we decide to remove *stand:release_pos_x* instead of *release_pos_x*.

```{r}
# Fit temporary model without stand:release_pos_x
fb_model_temp <- lm(whiff ~ 
                 stand + release_speed + release_spin_rate +
                 release_extension + release_pos_z + 
                 release_pos_x + plate_x + plate_z + 
                 ax + az + pfx_x + pfx_z + 
                 stand:ax + stand:plate_x,
               data = RHP_FB0, weights = count)

# coefficients, correlation, and statistical significance
summary(fb_model_temp)

# Perform VIF test
vif(fb_model_temp)
```

As anticipated, without *stand:release_pos_x*, *release_pos_x* has an exceptionally low p-value, and a substantially increased coefficient estimate. The inflation factor of *stand* is also decreased significantly.

Predictor variable *ax* relamins to have an exceptionally high p-value.

Before conducting this research, we anticipated *ax* to be one of the most important predictors—in order for the ball to miss the bat, it needs to accelerate away from it.

From the VIF test, we can see that *ax*, *az*, *pfx_x*, and *pfx_z* stand out as highly inflated variables.

On the basis of this, we hypothesize that *pfx_x* is collinear with *ax*, and *pfx_z* is collinear with *az*. The reasoning is that horizontal/vertical acceleration causes horizontal_vertical movement.

```{r}
# Fit temporary model without pfx_x, pfx_z
fb_model_temp <- lm(whiff ~ 
                 stand + release_speed + release_spin_rate +
                 release_extension + release_pos_z + release_pos_x + plate_z +
                 plate_x + ax + az +
                 release_pos_x + stand:plate_x + stand:ax,
               data = RHP_FB0, weights = count)

# coefficients, correlation, and statistical significance
summary(fb_model_temp)

# Perform VIF test
vif(fb_model_temp)

# Perform Partial F test
anova(fb_model_temp, fb_model)
```

After removing *pfx_x* and *pfx_z*, we can observe that the p-values of *ax* and *az* have decreased. Also, the VIF of *ax* and *az* have also decreased significantly.

Hence, we decide to remove *pfx_x* and *pfx_z*. Although the partial F test shows statistically significant difference in residuals, we believe it is not substantial enough to justify the multicollinearity; there is no change in Adjusted R-squared (and R-squared) to the third decimal point, in the new model.

## Task 4: Update Linear Model

```{r}
# Rename
fb_model_updated <- fb_model_temp

# Remove old model to clean up workspace
rm(fb_model)
rm(fb_model_temp)
```

```{r}
# Residual vs Fitted plot
y_hat <- fitted(fb_model_updated)
e_hat <- resid(fb_model_updated)
plot(x = y_hat, y = e_hat, main = "Residual vs Fitted", xlab = "Fitted", ylab = "Residuals")

# Create QQ plot
qqnorm(e_hat)
qqline(e_hat)

# Create condition 2 plot
pairs(RHP_FB0[, setdiff(colnames(RHP_FB0)[4:15], c("release_pos_x", "pfx_x", "pfx_z"))])

```

Description

Description

Description

Description

Description

## Task 5: Predictions

-   Create a whiff rate prediction for each individual pitch
-   Aggregate the predictions, and the actual results for each pitcher, each year
-   Filter pitchers with too few results (We decided to use count \> 50 as that removes all the whiff = 0; In MLB, pitchers regularily throw 1000+ pitches throughout the season.)

```{r}
# Create predictions for whiff rate
Predict_RHP_FB0 <- RHP_FB0 |>
  ungroup() |>
  mutate(est_whiff = predict(fb_model_updated, newdata = RHP_FB0))

# Group predictions by player and year
Final <- Predict_RHP_FB0 |>
  group_by(player_name, game_year) |>
  summarise(new_count = sum(count),
            whiff = sum(whiff * count) / sum(count),
            est_whiff = sum(est_whiff * count) / sum(count))

# Clean up workspace
remove(RHP_FB0)
remove(Predict_RHP_FB0)
remove(fb_model_updated)
```

## Task 6: Final Diagnostics and Analysis

-   Use a linear model between prediction and actual to analyze the results
    -   We will use a weighted linear model as some pitchers have thrown more pitches than others

```{r}
# Fit a linear model between predicted and actual whiff rates
Final_lm <- lm(whiff ~ est_whiff, data = Final, weights = new_count)

# Summary of the final model
summary(Final_lm)

# Plot predicted vs actual whiff rate
ggplot(Final, aes(x = est_whiff, y = whiff)) +
  geom_point(color = "blue", alpha = 0.6, size = 2) +
  geom_abline(intercept = 0, slope = 1, linetype = "solid", color = "black") +
  labs(
    title = "Predicted vs Actual Whiff Rate for Pitchers",
    x = "Estimated Whiff Rate (Predicted)",
    y = "Actual Whiff Rate"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10)),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank()
  )
```

```{r}
# Residual vs Fitted plot
y_hat <- fitted(Final_lm)
e_hat <- resid(Final_lm)
plot(x = y_hat, y = e_hat, main = "Residual vs Fitted", xlab = "Fitted", ylab = "Residuals")

# Create QQ plot
qqnorm(e_hat)
qqline(e_hat)
```